---
title: "Overview of the `wcde` package"
output:
  html_document:
    fig_caption: false
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Overview of wcde}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `wcde` package allows for R users to easily download data from the [Wittgenstein Centre Human Capital Data Explorer](http://dataexplorer.wittgensteincentre.org/wcde-v2/) as well as containing a number of summary data sets and helpful functions for working with education specific demographic data.

# Getting data into R

The `get_wcde()` function can be used to download data from the Wittgenstein Centre Human Capital Data Explorer. It requires three user inputs

- `indicator`: a short code for the indicator of interest
- `scenario`: a number referring to a SSP narrative, by default 2 is used (for SSP2) 
- `country_code` (or `country_name`): corresponding to the country of interest 

```{r, messages = FALSE, message=FALSE}
library(wcde)
# download education specific tfr data
get_wcde(indicator = "etfr", 
         country_name = c("Brazil", "Albania"))

# download education specific survivorship rates
get_wcde(indicator = "eassr", 
         country_name = c("Niger", "Korea"))
```

## Indicator codes

The indicator input must match the short code from the indicator table. The `find_indicator()` function can be used to look up short codes (given in the first column) from the `wic_indicators` data frame:

```{r}
find_indicator(x = "tfr")
```

## Temporal coverage 

By default, `get_wdce()` returns data for all years or available periods or years. The `filter()` function in dplyr can be used to filter data for specific years or periods, for example:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
get_wcde(indicator = "e0", 
         country_name = c("Japan", "Australia")) %>%
  filter(period == "2015-2020")

get_wcde(indicator = "sexratio", 
         country_name = c("China", "South Korea")) %>%
  filter(year == 2020)
```

Past data is only available for selected indicators. These can be viewed using the `past` indicator column: 

```{r}
wic_indicators %>%
  filter(past) %>%
  select(1:2)
```

The filter function can also be used to filter specific indicators to specific age, sex or education groups

```{r, messages = FALSE, message=FALSE}
get_wcde(indicator = "sexratio", 
         country_name = c("China", "South Korea")) %>%
  filter(year == 2020, 
         age == "All")
```


## Country names and codes

Country names are guessed using the [countrycode](https://cran.r-project.org/web/packages/countrycode/index.html) package. 

The `get_wcde()` accepts ISO alpha_numeric codes for countries:

```{r, messages = FALSE, message=FALSE}
get_wcde(indicator = "tfr", 
         country_name = c("U.A.E", "Espania", "Österreich"))

get_wcde(indicator = "etfr", country_code = c(44, 100))
```
A full list of available countries and region aggregates can be found in the `wic_locations` data frame. 

```{r}
wic_locations
```


## Scenarios

By default `get_data()` returns data for Medium (SSP2) scenario. Results for different SSP scenarios can be returned by passing a different (or multiple) scenario values to the `scenario` argument in `get_data()`. 

```{r, messages = FALSE, message=FALSE}
get_wcde(indicator = "growth", 
         country_name = c("India", "China"), 
         scenario = 1:5) %>%
  filter(period == "2095-2100")
```

Set `include_scenario_names = TRUE` to include a columns with the full names of the scenarios

```{r, messages = FALSE, message=FALSE}
get_wcde(indicator = "tfr", 
         country_name = c("Kenya", "Nigeria", "Algeria"), scenario = c(2, 1, 5), 
         include_scenario_names = TRUE) %>%
  filter(period == "2045-2050")
```

Additional details of the pathways for each scenario numeric code can be found in the `wic_scenarios` object. Further background and links to the corresponding literature are provided in the [Data Explorer](http://dataexplorer.wittgensteincentre.org/wcde-v2/) 

```{r}
wic_scenarios
```



## All countries data 

Data for all countries can be obtained by not setting `country_name` or `country_code`

```{r, messages = FALSE, message=FALSE}
get_wcde(indicator = "mage")
```

## Multiple indicators

The `get_wdce()` function needs to be called multiple times to download multiple indicators. This can be done using the `map()` function in [`purrr`](https://cran.r-project.org/web/packages/purrr/index.html)

```{r, messages = FALSE, message=FALSE}
mi <- tibble(ind = c("odr", "nirate", "ggapedu25")) %>%
  mutate(d = map(.x = ind, .f = ~get_wcde(indicator = .x)))
mi

mi %>%
  filter(ind == "odr") %>%
  select(-ind) %>%
  unnest(cols = d)

mi %>%
  filter(ind == "nirate") %>%
  select(-ind) %>%
  unnest(cols = d)

mi %>%
  filter(ind == "ggapedu25") %>%
  select(-ind) %>%
  unnest(cols = d)
```


# Working with population data

The education population data in the data explorer, obtained by setting `indicator = "epop"` in `get_wcde()`, provide results by up to three different education categorizations (4, 6 and 8 education groups). 

```{r, messages = FALSE, message=FALSE}
d <- get_wcde(indicator = "epop", country_code = 900)
d
```

As the data frame contains multiple groupings, the `edu_group_sum()` function can be used to provide education specific population data. Users can specify the education groupings by setting the `n` argument to 4, 6 or 8.

```{r, warning=FALSE, message=FALSE}
d %>% 
  edu_group_sum(n = 4) %>%
  filter(year == 2020)

d %>% 
  edu_group_sum(n = 6) %>%
  filter(year == 2020,
         age == "30--34")
```

# Population pyramids

Create population pyramids by setting male population values to negative equivalent to allow for divergent columns from the y axis.

```{r}
w <- d %>% 
  edu_group_sum(n = 4) %>%
  filter(year == 2020) %>%
  mutate(pop = ifelse(test = sex == "Male", yes = -epop, no = epop),
         pop = pop/1e3) 
w
```

Use standard ggplot code to create population pyramid with

-   `scale_x_symmetric()` from the [lemon](https://cran.r-project.org/web/packages/lemon/index.html) package to allow for equal male and female x-axis
-   fill colours set to the `wic_col4` object in the wcde package which contains the names of the colours used in the Wittgenstein Centre Human Capital Data Explorer Data Explorer.

Note `wic_col6` and `wic_col8` objects also exist for equivalent plots of population data objects with corresponding numbers of categories of education.

```{r, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}
library(lemon)

ggplot(data = w, 
       mapping = aes(x = pop, y = age, fill = fct_rev(education))) +
  geom_col() +
  geom_vline(xintercept = 0, colour = "black") + 
  scale_x_symmetric(labels = abs) +
  scale_fill_manual(values = wic_col4, name = "Education") +
  theme_bw()
```

Add male and female labels on the x-axis by 

- Creating a facet plot with the strips on the bottom with transparent backgrounds and no space between.
- Set the x axis to have zero expansion beyond the values in the data allowing the two sides of the pyramids to meet.
- Add a `geom_blank()` to allow for equal x-axis and additional space at the end of largest columns.

```{r, fig.width=6, fig.height=6}
w <- w %>%
  mutate(pop_max = ifelse(sex == "Male", -max(pop), max(pop)))

ggplot(data = w, 
       mapping = aes(x = pop, y = age, fill = fct_rev(education))) +
  geom_col() +
  geom_vline(xintercept = 0, colour = "black") +
  geom_blank(mapping = aes(x = pop_max * 1.05)) +
  scale_x_continuous(labels = abs, expand = c(0, 0)) +
  facet_wrap(facets = "sex", scales = "free_x", strip.position = "bottom") +
  scale_fill_manual(values = wic_col4, name = "Education") +
  theme(panel.spacing.x = unit(0, "pt"),
        strip.placement = "outside",
        strip.background = element_rect(fill = "transparent"),
        strip.text.x = element_text(margin = margin( b = 0, t = 0)))
```
